\Large Feature and order manipulations in a free recall task affect memory for current and future lists

### 2023-11-14 paper/main.tex
updated stats in main text

To obtain a clean preliminary estimate of the consequences on memory of
randomly varying the font colors and locations of presented words (versus
holding the font color fixed at black, and holding the display locations fixed
at the center of the display) we compared participants' performance on the
\textit{feature rich} and \textit{reduced} experimental conditions (see
\textit{Random conditions}, Fig.~\dynamicsRandom). In the feature rich
condition the words' colors and locations varied randomly across words, and in
the reduced condition words were always presented in black, at the center of
the display. Aggregating across all lists for each participant, we found no
difference in recall accuracy (i.e., the proportions of correctly recalled
words) for feature rich versus reduced lists ($t(126) = -0.290, p = 0.772$).
However, participants in the feature rich condition clustered their recalls
substantially more along every dimension we examined (temporal clustering:
$t(126) = 10.624, p < 0.001$; semantic category clustering: $t(126) = 10.077, p
< 0.001$; size clustering: $t(126) = 11.829, p < 0.001$; word length
clustering: $t(126) = 10.639, p < 0.001$; first letter clustering: $t(126) =
7.775, p < 0.001$; see \textit{Permutation-corrected feature clustering scores}
for more information about how we quantified each participant's clustering
tendencies.) Taken together, these comparisons suggest that adding new features
changes how participants organize their memories of studied words, even when
those new features are independent of the words themselves and even when the
new features vary randomly across words. We found no evidence that those
additional uninformative features were distracting (in terms of their impact on
memory performance), but they did affect participants' recall dynamics
(measured via their clustering scores).

We also wondered whether adding these incidental visual features to later lists
(after the participants had already studied impoverished lists), or removing
the visual features from later lists (after the participants had already
studied visually diverse lists) might affect memory performance. In other
words, we sought to test for potential effects of changing the ``richness'' of
participants' experiences over time. All participants studied and recalled a
total of 16 lists; we defined \textit{early} lists as the first eight lists and
\textit{late} lists as the last eight lists each participant encountered. To
help interpret our results, we compared participants' memories on early versus
late lists in the above feature rich and reduced conditions. Participants in
both conditions remembered more words on early versus late lists (feature rich:
$t(66) = 4.553, p < 0.001$; reduced: $t(60) = 2.434, p = 0.018$). Participants
in the feature rich (but not reduced) conditions exhibited more temporal
clustering on early versus late lists (feature rich: $t(66) = 2.318, p =
0.024$; reduced: $t(60) = 0.929, p = 0.357$). And participants in both
conditions exhibited more semantic (category and size) clustering on early
versus late lists (feature rich, category: $t(66) = 3.805, p < 0.001$; feature
rich, size: $t(66) = 2.190, p = 0.032$; reduced, category: $t(60) = 2.856, p =
0.006$; reduced, size: $t(60) = 2.947, p = 0.005$). Participants in the reduced
(but not feature rich) conditions exhibited more lexicographic clustering on
early versus late lists (feature rich, word length: $t(66) = 0.161, p = 0.872$;
feature rich, first letter: $t(66) = 0.410, p = 0.683$; reduced, word length:
$t(60) = 3.528, p = 0.001$; reduced, first letter: $t(60) = 2.275, p = 0.026$).
Taken together, these comparisons suggest that even when the presence or
absence of incidental visual features is stable across lists, participants
still exhibit some differences in their performance and memory organization
tendencies for early versus late lists.

With these differences in mind, we next compared participants' memories on
early versus late lists for two additional experimental conditions (see
\textit{Random conditions}, Fig.~\dynamicsRandom). In a \textit{reduced
(early)} condition, we held the visual features constant on early lists, but
allowed them to vary randomly on late lists. In a \textit{reduced (late)}
condition, we allowed the visual features to vary randomly on early lists, but
held them constant on late lists. Given our above findings that (a)
participants tended to exhibit stronger clustering effects on feature rich
(versus reduced) lists, and (b) participants tended to remember more words and
exhibit stronger clustering effects on early (versus late) lists, we expected
these early versus late differences to be enhanced in the reduced (early)
condition and diminished in the reduced (late) condition. However, to our
surprise, participants in \textit{neither} condition exhibited reliable early
versus late differences in accuracy (reduced (early): $t(41) = 1.499, p =
0.141$; reduced (late): $t(40) = 1.462, p = 0.152$), temporal clustering
(reduced (early): $t(41) = 0.998, p = 0.324$; reduced (late): $t(40) = 1.099, p
= 0.278$), nor feature-based clustering (reduced (early), category: $t(41) =
0.753, p = 0.456$; reduced (early), size: $t(41) = 0.721, p = 0.475$; reduced
(early), length: $t(41) = 0.493, p = 0.625$; reduced (early), first letter:
$t(41) = 0.780, p = 0.440$; reduced (late), category: $t(40) = -0.086, p =
0.932$; reduced (late), size: $t(40) = 0.746, p = 0.460$; reduced (late),
length: $t(40) = 1.476, p = 0.148$; reduced (late), first letter: $t(40) =
0.966, p = 0.340$). We hypothesized that adding or removing the variability in
the visual features was acting as a sort of ``event boundary'' between early
and late lists~\citep[e.g., ][]{ClewEtal19,RadvZack17,RadvCope06}. In prior
work, we (and others) have found that memories formed just after event
boundaries can be enhanced~\citep[e.g., due to less contextual interference
between pre- and post-boundary items;][]{MannEtal16, PettEtal16, GoldEtal17,
FlorEtal17}.

We found that \textit{adding} incidental visual features on later lists that
had not been present on early lists (as in the reduced (early) condition)
served to enhance recall performance relative to conditions where all lists had
the same blends of features (accuracy for feature rich versus reduced (early):
$t(107) = -2.230, p = 0.028$; reduced versus reduced (early): $t(101) = -2.045,
p = 0.043$; also see Fig.~\accuracyByList A). However, \textit{subtracting}
irrelevant visual features on later lists that \textit{had} been present on
early lists (as in the reduced (late) condition) did not appear to impact
recall performance (accuracy for feature rich versus reduced (late): $t(106) =
-0.638, p = 0.525$; reduced versus reduced (late): $t(100) = -0.407, p =
0.685$). These comparisons suggest that recall accuracy has a directional
component: accuracy is affected differently by removing features later
that had been present earlier versus adding features later that had
\textit{not} been present earlier. In contrast, we found that participants
exhibited more temporal and feature-based clustering when we added incidental
visual features to \textit{any} lists (comparisons of clustering on feature
rich versus reduced lists are reported above; temporal clustering in reduced
versus reduced (early) and reduced versus reduced (late) conditions: $t$s $\leq
-9.780$, $p$s $< 0.001$; feature-based clustering in reduced versus reduced
(early) and reduced versus reduced (late) conditions: $t$s $\leq -5.443$, $p$s
$< 0.001$). Temporal and feature-based clustering were not reliably different
in the feature rich, reduced (early), and reduced (late) conditions (temporal
clustering in feature rich versus reduced (early) and feature rich versus
reduced (late) conditions: $t$s $\geq -1.434$, $p$s $\geq 0.154$; feature-based
clustering in feature rich versus reduced (early) and feature rich versus
reduced (late) conditions: $t$s $\geq -1.359$, $p$s $> 0.177$).

@@

early feature rich lists; $t(95) = 3.034, p = 0.003$). Participants in the
color-ordered condition also showed a trending increase in memory performance
on early lists (again, relative to early feature rich lists: $t(96) = 1.850, p
= 0.067$; Fig.~\ref{fig:barplots}A). Participants' performances on early lists
in all of the other order manipulation conditions were indistinguishable from
performance on the early feature rich lists ($\|t\|$s $< 1.013, p$s $> 0.314$).
Participants in both of the semantically ordered conditions exhibited stronger
temporal clustering on early lists (versus early feature rich lists; category:
$t(95) = 8.508, p < 0.001$; size: $t(95) = 2.429, p = 0.017$;
Fig.~\ref{fig:barplots}B). Participants in the length-ordered condition tended
to exhibit \textit{less} temporal clustering on early lists relative to early
feature rich lists ($t(95) = -1.666, p = 0.099$), whereas participants in the
first letter-ordered condition exhibited stronger temporal clustering on early
lists ($t(95) = 2.587, p = 0.011$). Participants in the visually ordered
conditions exhibited more similar performance on early lists, relative to early
feature rich lists (color: $t(96) = -1.064, p = 0.290$; we found a trending
enhancement for participants in the location-ordered condition: $t(95) = 1.682,
p = 0.096$). We also compared feature-based clustering on early lists across
the order manipulation and feature rich conditions. Since these results were

@@

condition (category: $t(125) = 2.524, p = 0.013$; size: $t(125) = 3.510, p =
0.001$), but showed no reliable differences in lexicographic (length: $t(125) =
0.539, p = 0.591$; first letter: $t(125) = -0.587, p = 0.558$) or visual
(color: $t(125) = -0.579, p = 0.564$; location: $t(125) = -0.346, p = 0.730$)
clustering. Similarly, participants in the lexicographically ordered conditions
exhibited stronger (relative to feature rich participants) lexicographic
clustering (length: $t(125) = 3.426, p = 0.001$; first letter: $t(125) = 3.236,
p = 0.002$) on early lists, but showed no reliable differences in semantic
(category: $t(125) = -1.078, p = 0.283$; size: $t(125) = -0.310, p = 0.757$) or
visual (color: $t(125) = -0.209, p = 0.835$; location: $t(125) = -0.004, p =
0.997$) clustering. And participants in the visually ordered conditions
exhibited stronger visual clustering (again, relative to feature rich
participants, and on early lists; color: $t(126) = 2.099, p = 0.038$; location:
$t(126) = 4.392, p < 0.001$), but showed no reliable differences in semantic
(category: $t(126) = 0.204, p = 0.839$; size: $t(126) = -0.093, p = 0.926$) or
lexicographic (length: $t(126) = 0.714, p = 0.476$; first letter: $t(126) =
0.820, p = 0.414$) clustering. Taken together, these order manipulation results
suggest several broad patterns (Figs.~\ref{fig:recall-dynamics}A,
\ref{fig:fingerprints}). First, most of the order manipulations we carried out
did \textit{not} reliably affect overall recall performance. Second, most of
the order manipulations increased participants' tendencies to temporally
cluster their recalls. Third, all of the order manipulations enhanced
participants' clustering of each condition's target feature (i.e., semantic
manipulations enhanced semantic clustering, lexicographic manipulations
enhanced lexicographic clustering, and visual manipulations enhanced visual
clustering; Fig.~\ref{fig:barplots}C) while leaving clustering along other
feature dimensions roughly unchanged (i.e., semantic manipulations did not
affect lexicographic or visual clustering, and so on). Although it is not
possible to fully separate feature versus temporal clustering when considering
sorted lists, we used a permutation-based procedure to identify the degree of
feature clustering over and above what could be accounted for by temporal
clustering alone (see \textit{Factoring out the effects of temporal
clustering}). When we carried out this analysis (Fig.~\ref{fig:barplots}D), we
found that participants exhibited more semantic clustering on semantically
sorted lists than on randomly ordered lists (category: $t(XXX) = XXX, p = XXX$;
size: $t(XXX) = XXX, p = XXX$), but the effects of the other order
manipulations could not reliably be separated from temporal clustering alone
($t$s $> XXX, p$s $> XXX$).

@@

The above comparisons between memory performance on early lists in the order
manipulation versus feature rich conditions highlight how sorted lists are
remembered differently from random lists. We also wondered how sorting lists
along each feature dimension influenced memory relative to sorting lists along
the other feature dimensions. Participants trended towards remembering early
lists that were sorted semantically better than lexicographically sorted lists
($t(118) = 1.936, p = 0.055$). Participants also remembered visually sorted
lists better than lexicographically sorted lists ($t(119) = 2.145, p = 0.034$).
However, participants showed no reliable differences in recall for semantically
versus visually sorted lists ($t(119) = 0.113, p = 0.910$). Participants
temporally clustered semantically sorted lists more strongly than either
lexicographically ($t(118) = 5.572, p < 0.001$) or visually ($t(119) = 6.215, p
< 0.001$) sorted lists, but did not show reliable differences in temporal
clustering on lexicographically versus visually sorted lists ($t(119) = 0.189,
p = 0.850$). Participants also showed reliably more semantic clustering on
semantically sorted lists than lexicographically (category: $t(118) = 3.492, p
= 0.001$, size: $t(118) = 3.972, p < 0.001$) or visually (category: $t(119) =
2.702, p = 0.008$, size: $t(119) = 4.230, p < 0.001$) sorted lists; more
lexicographic clustering on lexicographically sorted lists than semantically
(length: $t(118) = 3.112, p = 0.002$; first letter: $t(118) = 3.686, p <
0.001$) or visually (length: $t(119) = 3.024, p = 0.003$; first letter: $t(119)
= 2.644, p = 0.009$) sorted lists; and more visual clustering on visually
sorted lists than semantically (color: $t(119) = -2.659, p = 0.009$; location:
$t(119) = -4.604, p < 0.001$) or lexicographically (color: $t(119) = -2.366, p
= 0.020$; location: $t(119) = -4.265, p < 0.001$) sorted lists. In summary,
sorting lists by different features appeared to have slightly different effects
on overall memory performance and temporal clustering.  Participants also tended to
cluster their recalls along a given feature dimension more when the studied
lists were (versus were not) sorted along that dimension.

@@

At the group level, there seemed to be almost no lingering impact of sorting
early lists on memory for later lists. To simplify the presentation, we report
these null results in aggregate across the three feature groupings. Relative to
memory performance on late feature rich lists, participants' memory performance
in all six order manipulation conditions showed no reliable differences
(semantic: $t(125) = 0.487, p = 0.627$; lexicographic: $t(125) = 0.878, p =
0.382$; visual: $t(126) = 1.437, p = 0.153$). Nor did we observe any reliable
differences in temporal clustering on late lists (relative to late feature rich
lists; semantic: $t(125) = 0.146, p = 0.884$; lexicographic: $t(125) = 0.923, p
= 0.358$; visual: $t(126) = 0.525, p = 0.601$). Aside from a slightly increased
tendency for participants to cluster words by their length on late visual order
manipulation lists (more than late feature rich lists; $t(126) = 2.199, p =
0.030$), we observed no reliable differences in any type of feature clustering
on late order manipulation condition lists versus late feature rich lists
($\|t\|$s $\leq 1.234, p$s $\geq 0.220$).

@@

across conditions and participants (early: $r(179) = 0.537, p < 0.001$; late: $
r(179) = 0.492, p < 0.001$) as well as for each condition individually for
early ($r$s $\geq 0.386$, all $p$s $\leq 0.035$) and late ($r$s $\geq 0.462$,
all $p$s $\leq 0.010$) lists. We found no evidence of a condition-level trend;
for example, the conditions where participants tended to show stronger
clustering scores were not correlated with the conditions where participants
remembered more words (early: $r(4) = 0.526, p = 0.284$; late: $r(4) = -0.257,
p = 0.623$; see insets of Fig.~\ref{fig:clustering-scatterplots}A and B). We
observed carryover associations between feature clustering and recall
performance (Fig.~\ref{fig:clustering-scatterplots}C, D). Participants who
showed stronger feature clustering on early lists tended to recall more items
on late lists (across conditions: $r(179) = 0.492, p < 0.001$; all conditions

>>>>

To obtain a clean preliminary estimate of the consequences on memory of
randomly varying the font colors and locations of presented words (versus
holding the font color fixed at black, and holding the display locations fixed
at the center of the display) we compared participants' performance on the
\textit{feature rich} and \textit{reduced} experimental conditions (see
\textit{Random conditions}, Fig.~\dynamicsRandom). In the feature rich
condition the words' colors and locations varied randomly across words, and in
the reduced condition words were always presented in black, at the center of
the display. Aggregating across all lists for each participant, we found no
difference in recall accuracy (i.e., the proportions of correctly recalled
words) for feature rich versus reduced lists ($t(126) = -0.290, p = 0.772,
\mathrm{Cohen's}~d~(d) = -0.051,
\mathrm{bootstrap~estimated~95\%~confidence~interval}~(\mathrm{CI}) = [-2.387,
1.768]$). However, participants in the feature rich condition clustered their
recalls substantially more along every dimension we examined (temporal
clustering: $t(126) = 10.632, p < 0.001, d = 1.882, \mathrm{CI} = [7.786,
14.386]$; semantic category clustering: $t(126) = 10.148, p < 0.001, d = 1.796,
\mathrm{CI} = [7.324, 13.778] $; size clustering: $t(126) = 12.033, p < 0.001,
d = 2.129, \mathrm{CI} = [9.030, 15.918]$; word length clustering: $t(126) =
10.720, p < 0.001, d = 1.897, \mathrm{CI} = [7.442, 15.174]$; first letter
clustering: $t(126) = 6.679, p < 0.001, d = 1.182, \mathrm{CI} = [4.490,
9.611]$; see \textit{Permutation-corrected feature clustering scores} for more
information about how we quantified each participant's clustering tendencies.)
Taken together, these comparisons suggest that adding new features changes how
participants organize their memories of studied words, even when those new
features are independent of the words themselves and even when the new features
vary randomly across words. We found no evidence that those additional
uninformative features were distracting (in terms of their impact on memory
performance), but they did affect participants' recall dynamics (measured via
their clustering scores).

We also wondered whether adding these incidental visual features to later lists
(after the participants had already studied impoverished lists), or removing
the visual features from later lists (after the participants had already
studied visually diverse lists) might affect memory performance. In other
words, we sought to test for potential effects of changing the ``richness'' of
participants' experiences over time. All participants studied and recalled a
total of 16 lists; we defined \textit{early} lists as the first eight lists and
\textit{late} lists as the last eight lists each participant encountered. To
help interpret our results, we compared participants' memories on early versus
late lists in the above feature rich and reduced conditions. Participants in
both conditions remembered more words on early versus late lists (feature rich:
$t(66) = 4.553, p < 0.001, d = 0.233, \mathrm{CI} = [2.427, 7.262]$; reduced:
$t(60) = 2.434, p = 0.018, d = 0.134, \mathrm{CI} = [0.493, 4.910]$).
Participants in the feature rich (but not reduced) conditions exhibited more
temporal clustering on early versus late lists (feature rich: $t(66) = 2.268, p
= 0.027, d = 0.181, \mathrm{CI} = [0.437, 4.425]$; reduced: $t(60) = 0.986, p =
0.328, d = 0.061, \mathrm{CI} = [-0.897, 3.348]$). And participants in both
conditions tended to exhibit more semantic clustering on early versus late
lists (feature rich, category: $t(66) = 3.684, p < 0.001, d = 0.220,
\mathrm{CI} = [1.733, 5.732]$; feature rich, size: $t(66) = 1.629, p = 0.108, d
= 0.100, \mathrm{CI} = [-0.207, 3.905] $; reduced, category: $t(60) = 2.755, p
= 0.008, d = 0.177, \mathrm{CI} = [0.761, 5.189]$; reduced, size: $t(60) =
3.081, p = 0.003, d = 0.201, \mathrm{CI} = [1.210, 5.326]$). Participants in
the reduced (but not feature rich) conditions tended to exhibit more
lexicographic clustering on early versus late lists (feature rich, word length:
$t(66) = -0.100, p = 0.921, d = -0.010, \mathrm{CI} = [-2.217, 1.899]$; feature
rich, first letter: $t(66) = -0.412, p = 0.681, d = -0.045, \mathrm{CI} =
[-2.461, 1.645]$; reduced, word length: $t(60) = 3.762, p < 0.001, d = 0.261,
\mathrm{CI} = [1.604, 6.821]$; reduced, first letter: $t(60) = 1.721, p =
0.090, d = 0.175, \mathrm{CI} = [-0.138, 4.098]$). Taken together, these
comparisons suggest that even when the presence or absence of incidental visual
features is stable across lists, participants still exhibit some differences in
their performance and memory organization tendencies for early versus late
lists.

With these differences in mind, we next compared participants' memories on
early versus late lists for two additional experimental conditions (see
\textit{Random conditions}, Fig.~\dynamicsRandom). In a \textit{reduced
(early)} condition, we held the visual features constant on early lists, but
allowed them to vary randomly on late lists. In a \textit{reduced (late)}
condition, we allowed the visual features to vary randomly on early lists, but
held them constant on late lists. Given our above findings that (a)
participants tended to exhibit stronger clustering effects on feature rich
(versus reduced) lists, and (b) participants tended to remember more words and
exhibit stronger clustering effects on early (versus late) lists, we expected
these early versus late differences to be enhanced in the reduced (early)
condition and diminished in the reduced (late) condition. However, to our
surprise, participants in \textit{neither} condition exhibited reliable early
versus late differences in accuracy (reduced (early): $t(41) = 1.499, p =
0.141, d = 0.098, \mathrm{CI} = [-0.345, 3.579]$; reduced (late): $t(40) =
1.462, p = 0.152, d = 0.121, \mathrm{CI} = [-0.376, 2.993]$), temporal
clustering (reduced (early): $t(41) = 0.857, p = 0.396, d = 0.068, \mathrm{CI}
= [-1.012, 2.896]$; reduced (late): $t(40) = 1.244, p = 0.221, d = 0.128,
\mathrm{CI} = [-0.894, 3.088]$), nor feature-based clustering (reduced (early),
category: $t(41) = 0.707, p = 0.484, d = 0.068, \mathrm{CI} = [-1.314, 2.830]$;
reduced (early), size: $t(41) = 0.803, p = 0.427, d = 0.079, \mathrm{CI} =
[-1.142, 2.953]$; reduced (early), length: $t(41) = 0.461, p = 0.648, d =
0.060, \mathrm{CI} = [-1.545, 2.462]$; reduced (early), first letter: $t(41) =
0.781, p = 0.439, d = 0.101, \mathrm{CI} = [-1.039, 2.881]$; reduced (late),
category: $t(40) = -0.101, p = 0.920, d = -0.009, \mathrm{CI} = [-2.307,
1.776]$; reduced (late), size: $t(40) = 0.555, p = 0.582, d = 0.058,
\mathrm{CI} = [-1.444, 2.274] $; reduced (late), length: $t(40) = 1.482, p =
0.146, d = 0.126, \mathrm{CI} = [-0.444, 3.743]$; reduced (late), first letter:
$t(40) = -0.143, p = 0.887, d = -0.017, \mathrm{CI} = [-2.204, 1.830] $). We
hypothesized that adding or removing the variability in the visual features was
acting as a sort of ``event boundary'' between early and late
lists~\citep[e.g., ][]{ClewEtal19,RadvZack17,RadvCope06}. In prior work, we
(and others) have found that memories formed just after event boundaries can be
enhanced~\citep[e.g., due to less contextual interference between pre- and
post-boundary items;][]{MannEtal16, PettEtal16, GoldEtal17, FlorEtal17}.

We found that \textit{adding} incidental visual features on later lists that
had not been present on early lists (as in the reduced (early) condition)
served to enhance recall performance relative to conditions where all lists had
the same blends of features (accuracy for feature rich versus reduced (early):
$t(107) = -2.230, p = 0.028, d = -0.439, \mathrm{CI} = [-4.252, -0.229]$;
reduced versus reduced (early): $t(101) = -2.045, p = 0.043, d = -0.410,
\mathrm{CI} = [-3.826, 0.112]$; also see Fig.~\accuracyByList A). However,
\textit{subtracting} irrelevant visual features on later lists that
\textit{had} been present on early lists (as in the reduced (late) condition)
did not appear to impact recall performance (accuracy for feature rich versus
reduced (late): $t(106) = -0.638, p = 0.525, d = -0.126, \mathrm{CI} = [-2.720,
1.362]$; reduced versus reduced (late): $t(100) = -0.407, p = 0.685, d =
-0.082, \mathrm{CI} = [-2.477, 1.626]$). These comparisons suggest that recall
accuracy has a directional component: accuracy is affected differently by
removing features later that had been present earlier versus adding features
later that had \textit{not} been present earlier. In contrast, we found that
participants exhibited more temporal and feature-based clustering when we added
incidental visual features to \textit{any} lists (comparisons of clustering on
feature rich versus reduced lists are reported above; temporal clustering in
reduced versus reduced (early) and reduced versus reduced (late) conditions:
$t$s $\leq -9.885$, $p$s $< 0.001$; feature-based clustering in reduced versus
reduced (early) and reduced versus reduced (late) conditions: $t$s $\leq
-4.555$, $p$s $< 0.001$). Temporal and feature-based clustering were not
reliably different in the feature rich, reduced (early), and reduced (late)
conditions (temporal clustering in feature rich versus reduced (early) and
feature rich versus reduced (late) conditions: $t$s $\geq -1.379$, $p$s $\geq
0.171$; feature-based clustering in feature rich versus reduced (early) and
feature rich versus reduced (late) conditions: $|t|$s $\leq 1.441$, $p$s $\geq
0.153$).

@@

early feature rich lists; $t(95) = 3.034, p = 0.003, d = 0.667, \mathrm{CI} =
[1.048, 5.113]$). Participants in the color-ordered condition also showed a
trending increase in memory performance on early lists (again, relative to
early feature rich lists: $t(96) = 1.850, p = 0.067, d = 0.402, \mathrm{CI} =
[-0.010, 3.712]$; Fig.~\ref{fig:barplots}A). Participants' performances on
early lists in all of the other order manipulation conditions were
indistinguishable from performance on the early feature rich lists ($\|t\|$s
$\leq 1.013, p$s $\geq 0.314$). Participants in both of the semantically
ordered conditions exhibited stronger temporal clustering on early lists
(versus early feature rich lists; category: $t(95) = 8.813, p < 0.001, d =
1.936, \mathrm{CI} = [6.793, 11.751] $; size: $t(95) = 2.630, p = 0.010, d =
0.578, \mathrm{CI} = [0.831, 4.866]$; Fig.~\ref{fig:barplots}B). Participants
in the length-ordered condition tended to exhibit \textit{less} temporal
clustering on early lists relative to early feature rich lists ($t(95) =
-1.547, p = 0.125, d = -0.340, \mathrm{CI} = [-3.693, 0.341]$), whereas
participants in the first letter-ordered condition exhibited stronger temporal
clustering on early lists ($t(95) = 2.858, p = 0.005, d = 0.628, \mathrm{CI} =
[1.031, 4.886]$). Participants in the visually ordered conditions exhibited
more similar performance (accuracy) on early lists, relative to early feature
rich lists (we found a trending enhancement for participants in the
color-ordered condition: $t(96) = 1.850, p = 0.067, d = 0.402, \mathrm{CI} =
[-0.010, 3.712]$; location: $t(95) = 0.043, p = 0.966, d = 0.010, \mathrm{CI} =
[-1.598, 1.729]$). Participants in the visually ordered conditions also showed
similar temporal clustering on early lists, relative to early feature rich
lists (color: $t(96) = -1.339, p = 0.184, d = -0.291, \mathrm{CI} = [-3.238,
0.394]$, we found a trending increase for participants in the location-ordered
condition: $t(95) = 1.705, p = 0.092, d = 0.374, \mathrm{CI} = [-0.155,
3.521]$). We also compared feature-based clustering on early lists across the
order manipulation and feature rich conditions. Since these results were

@@

condition (category: $t(125) = 2.722, p = 0.007, d = 0.484, \mathrm{CI} =
[0.827, 4.932]$; size: $t(125) = 3.866, p < 0.001, d = 0.687, \mathrm{CI} =
[2.020, 5.983]$), but showed no reliable differences in lexicographic (length:
$t(125) = 0.521, p = 0.603, d = 0.093, \mathrm{CI} = [-1.311, 2.333]$; first
letter: $t(125) = -0.842, p = 0.401, d = -0.150, \mathrm{CI} = [-2.825,
1.095]$) or visual (color: $t(125) = -0.650, p = 0.517, d = -0.116, \mathrm{CI}
= [-2.680, 1.249]$; location: $t(125) = -0.251, p = 0.802, d = -0.045,
\mathrm{CI} = [-2.257, 1.524]$) clustering. Similarly, participants in the
lexicographically ordered conditions exhibited stronger (relative to feature
rich participants) lexicographic clustering (length: $t(125) = 3.682, p <
0.001, d = 0.655, \mathrm{CI} = [1.890, 5.569]$; first letter: $t(125) = 5.134,
p < 0.001, d = 0.912, \mathrm{CI} = [3.251, 7.258]$) on early lists, but showed
no reliable differences in semantic (category: $t(125) = -1.040, p = 0.301, d =
-0.185, \mathrm{CI} = [-3.095, 1.092]$; size: $t(125) = 0.006, p = 0.995, d =
0.001, \mathrm{CI} = [-1.933, 1.952]$) or visual (color: $t(125) = 0.092, p =
0.927, d = 0.016, \mathrm{CI} = [-1.834, 1.867]$; location: $t(125) = 0.407, p
= 0.685, d = 0.072, \mathrm{CI} = [-1.655, 2.463]$) clustering. And
participants in the visually ordered conditions exhibited stronger visual
clustering (again, relative to feature rich participants, and on early lists;
color: $t(126) = 2.022, p = 0.045, d = 0.358, \mathrm{CI} = [0.056, 3.965]$;
location: $t(126) = 4.390, p < 0.001, d = 0.777, \mathrm{CI} = [2.730, 6.199]
$), but showed no reliable differences in semantic (category: $t(126) = 0.012,
p = 0.991, d = 0.002, \mathrm{CI} = [-1.988, 1.871]$; size: $t(126) = -0.104, p
= 0.917, d = -0.018, \mathrm{CI} = [-2.166, 1.847] $) or lexicographic (length:
$t(126) = 0.592, p = 0.555, d = 0.105, \mathrm{CI} = [-1.361, 2.420]$; first
letter: $t(126) = 0.040, p = 0.968, d = 0.007, \mathrm{CI} = [-1.791, 1.863]$)
clustering. Taken together, these order manipulation results suggest several
broad patterns (Figs.~\ref{fig:recall-dynamics}A, \ref{fig:fingerprints}).
First, most of the order manipulations we carried out did \textit{not} reliably
affect overall recall performance. Second, most of the order manipulations
increased participants' tendencies to temporally cluster their recalls. Third,
all of the order manipulations enhanced participants' clustering of each
condition's target feature (i.e., semantic manipulations enhanced semantic
clustering, lexicographic manipulations enhanced lexicographic clustering, and
visual manipulations enhanced visual clustering; Fig.~\ref{fig:barplots}C)
while leaving clustering along other feature dimensions roughly unchanged
(i.e., semantic manipulations did not affect lexicographic or visual
clustering, and so on). Although it is not possible to fully separate feature
versus temporal clustering when considering sorted lists, we used a
permutation-based procedure to identify the degree of feature clustering over
and above what could be accounted for by temporal clustering alone (see
\textit{Factoring out the effects of temporal clustering}). When we carried out
this analysis (Fig.~\ref{fig:barplots}D), we found that participants exhibited
more semantic clustering on semantically sorted lists than on randomly ordered
lists, but the effects of the other order manipulations could not reliably be
separated from temporal clustering alone (reliable comparisons are reported in
the figure).

@@

The above comparisons between memory performance on early lists in the order
manipulation versus feature rich conditions highlight how sorted lists are
remembered differently from random lists. We also wondered how sorting lists
along each feature dimension influenced memory relative to sorting lists along
the other feature dimensions. Participants trended towards remembering early
lists that were sorted semantically better than lexicographically sorted lists
($t(118) = 1.936, p = 0.055, d = 0.353, \mathrm{CI} = [0.057, 3.916]$).
Participants also remembered visually sorted lists better than
lexicographically sorted lists ($t(119) = 2.145, p = 0.034, d = 0.390,
\mathrm{CI} = [0.208, 4.254]$). However, participants showed no reliable
differences in recall for semantically versus visually sorted lists ($t(119) =
0.113, p = 0.910, d = 0.021, \mathrm{CI} = [-1.987, 2.097]$). Participants
temporally clustered semantically sorted lists more strongly than either
lexicographically ($t(118) = 5.620, p < 0.001, d = 1.026, \mathrm{CI} = [3.486,
8.010]$) or visually ($t(119) = 6.613, p < 0.001, d = 1.202, \mathrm{CI} =
[4.481, 9.464]$) sorted lists, but did not show reliable differences in
temporal clustering on lexicographically versus visually sorted lists ($t(119)
= 0.589, p = 0.557, d = 0.107, \mathrm{CI} = [-1.336, 2.539]$). Participants
also showed reliably more semantic clustering on semantically sorted lists than
lexicographically (category: $t(118) = 3.667, p < 0.001, d = 0.670, \mathrm{CI}
= [1.822, 5.942]$, size: $t(118) = 3.972, p < 0.001$) or visually (category:
$t(119) = 2.702, p = 0.008$, size: $t(118) = 4.043, p < 0.001, d = 0.738,
\mathrm{CI} = [2.145, 6.296]$) sorted lists; more lexicographic clustering on
lexicographically sorted lists than semantically (length: $t(118) = 3.390, p <
0.001, d = 0.619, \mathrm{CI} = [1.499, 5.661]$; first letter: $t(118) = 5.705,
p < 0.001, d = 1.042, \mathrm{CI} = [3.841, 7.790]$) or visually (length:
$t(119) = 3.399, p < 0.001, d = 0.618, \mathrm{CI} = [1.500, 5.527]$; first
letter: $t(119) = 4.859, p < 0.001, d = 0.883, \mathrm{CI} = [2.860, 6.849]$)
sorted lists; and more visual clustering on visually sorted lists than
semantically (color: $t(119) = 2.673, p = 0.009, d = 0.486, \mathrm{CI} =
[0.848, 4.567]$; location: $t(119) = 4.499, p < 0.001, d = 0.818, \mathrm{CI} =
[2.721, 6.399]$) or lexicographically (color: $t(119) = 1.988, p = 0.049, d =
0.361, \mathrm{CI} = [0.102, 3.894]$; location: $t(119) = 3.966, p < 0.001, d =
0.721, \mathrm{CI} = [2.099, 5.862]$) sorted lists. In summary, sorting lists
by different features appeared to have slightly different effects on overall
memory performance and temporal clustering. Participants also tended to cluster
their recalls along a given feature dimension more when the studied lists were
(versus were not) sorted along that dimension.

@@

At the group level, there seemed to be almost no lingering impact of sorting
early lists on memory for later lists. To simplify the presentation, we report
these null results in aggregate across the three feature groupings. Relative to
memory performance on late feature rich lists, participants' memory performance
in all six order manipulation conditions showed no reliable differences
(semantic: $t(125) = 0.487, p = 0.627, d = 0.087, \mathrm{CI} = [-1.661,
2.323]$; lexicographic: $t(125) = 0.878, p = 0.382, d = 0.156, \mathrm{CI} =
[-1.226, 3.044]$; visual: $t(126) = 1.437, p = 0.153, d = 0.254, \mathrm{CI} =
[-0.447, 3.519]$). Nor did we observe any reliable differences in temporal
clustering on late lists (relative to late feature rich lists; semantic:
$t(125) = 0.157, p = 0.875, d = 0.028, \mathrm{CI} = [-1.859, 1.974]$;
lexicographic: $t(125) = 0.998, p = 0.320, d = 0.177, \mathrm{CI} = [-0.902,
2.920]$; visual: $t(126) = 0.548, p = 0.585, d = 0.097, \mathrm{CI} = [-1.450,
2.365]$). Aside from a slightly increased tendency for participants to cluster
words by their length on late visual order manipulation lists (more than late
feature rich lists; $t(126) = 2.005, p = 0.047, d = 0.355, \mathrm{CI} =
[0.211, 3.722]$), we observed no reliable differences in any type of feature
clustering on late order manipulation condition lists versus late feature rich
lists ($\|t\|$s $\leq 1.124, p$s $\geq 0.263$).

@@

across conditions and participants (early: $r(179) = 0.492, p < 0.001,
\mathrm{CI} = [0.352, 0.606]$; late: $r(179) = 0.403, p < 0.001, \mathrm{CI} =
[0.271, 0.517]$) as well as for each condition individually for early ($r$s
$\geq 0.331$, all $p$s $\leq 0.069$) and late ($r$s $\geq 0.404$, all $p$s
$\leq 0.027$) lists. We found no evidence of a condition-level trend; for
example, the conditions where participants tended to show stronger clustering
scores were not correlated with the conditions where participants remembered
more words (early: $r(4) = 0.511, p = 0.300, \mathrm{CI} = [-0.999, 0.996]$;
late: $r(4) = -0.304, p = 0.559, \mathrm{CI} = [-0.833, 0.748]$; see insets of
Fig.~\ref{fig:clustering-scatterplots}A and B). We observed carryover
associations between feature clustering and recall performance
